---
title: "Class 4: Static maps"
author: "Andrew Ba Tran"
output: 
  learnr::tutorial:
      theme: lumen
      highlight: espresso
      progressive: true
      allow_skip: false
      includes:
        before_body: _navbar.html
runtime: shiny_prerendered
# Do not index/display tutorial by setting `private: true`
private: true
description: >
  Making static choropleth maps
---


```{css, echo=FALSE}
.pageContent {
padding-top: 64px }
```

```{r setup, include=FALSE}
packages <- c("tidyverse", "lubridate", "rvest", "httr", "remotes",
              "tigris", "sf", "cowplot", "gridGraphics", "biscale",
              "ggspatial", "viridis")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos = "https://cran.us.r-project.org")  
}

#remotes::install_github("rstudio/gradethis", upgrade="always", quiet=TRUE)
#remotes::install_github("rstudio/learnr", upgrade="always", quiet=TRUE)

library(tidyverse)
library(learnr)
library(gradethis)
library(lubridate)
library(sf)
library(tigris)
library(cowplot)
library(biscale)
library(gridGraphics)
library(ggspatial)

us_states <- states(cb = TRUE, resolution = "20m") %>%
     shift_geometry()


us_counties <- counties(cb = TRUE, resolution = "20m") %>%
     shift_geometry()

df <- read_csv("https://www.fema.gov/api/open/v2/DisasterDeclarationsSummaries.csv")
df_new <- df %>% 
  mutate(GEOID=str_c(fipsStateCode, fipsCountyCode))

county_pop <- read_csv("data/county_population.csv")

fema_counties <- left_join(df_new, county_pop, by="GEOID")

fema_counties_count <- fema_counties %>%
  filter(incidentType!="Biological") %>% 
  group_by(GEOID) %>% 
  summarize(disasters=n())

fema_map <- left_join(us_counties, fema_counties_count)
#fema stuff


disaster_count <- fema_counties %>% 
  filter(incidentType=="Hurricane") %>% 
  mutate(year=year(incidentBeginDate)) %>% 
  filter(year %in% c(2017:2022)) %>% 
  group_by(GEOID) %>% 
  summarize(total=n())

disaster_counties <- left_join(us_counties, disaster_count) %>% 
  select(-NAME)
disaster_counties <- left_join(disaster_counties, county_pop)

bi_df <- bi_class(disaster_counties, x = estimate, y = total, style = "quantile", dim = 3)

map <- ggplot() +
  geom_sf(data = bi_df, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink", dim = 3) +
  labs(
    title = "Hurricanes and population"
  ) +
  bi_theme()

  
legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Higher population ",
                    ylab = "Higher disasters ",
                    size = 8)

legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Higher population ",
                    ylab = "Higher disasters ",
                    size = 8)

```


```{css, echo=FALSE}
.pageContent {
padding-top: 64px }

.cell-left {
  text-align: left;
}

.cell-right {
  text-align: right;
}

.cell-center {
  text-align: center;
}

table {
    margin: auto;
    border-top: 1px solid rgb(102, 102, 102);
    border-bottom: 1px solid rgb(102, 102, 102);
    display: table;
    border-collapse: separate;
    box-sizing: border-box;
    border-spacing: 2px;
    border-color: grey;
    padding-bottom:5px;
}
```

<span style="color:white">welcome to class!</span>


## Intro to sf


Builds on top of ggplot


## Get shapefiles

If you had the shapefiles locally, you could import it in using `st_read()`.

But because this is R, someone's already come up with a package to download the most-common ones from the cloud.


## Tigris

The Census has a server of all the most common boundary files.

We're going to work with the **tigris** package to download the shapefiles for counties and work with FEMA data!.

TIGRIS COMMANDS LIST GOES HERE


```{r tigris, exercise=TRUE}
library(tigris)
us_states <- states(cb = TRUE, resolution = "20m") %>%
     shift_geometry()

glimpse(us_states)
```

## sf



```{r tigris2, exercise=TRUE}

ggplot(us_states) + geom_sf()

```

```{r tigris4, exercise=TRUE}

us_counties <- counties(cb = TRUE, resolution = "20m") %>%
     shift_geometry()

glimpse(us_counties)

```

```{r tigris5, exercise=TRUE}
ggplot(us_counties) + 
  geom_sf() +
  theme_void()
```



```{r tigris6, exercise=TRUE}
df <- read_csv("https://www.fema.gov/api/open/v2/DisasterDeclarationsSummaries.csv")
df_new <- df %>% 
  mutate(GEOID=str_c(fipsStateCode, fipsCountyCode))

county_pop <- read_csv("data/county_population.csv")

fema_counties <- left_join(df_new, county_pop, by="GEOID")

glimpse(fema_counties)
```


```{r tigris7, exercise=TRUE}
fema_counties_count <- fema_counties %>%
  filter(incidentType!="Biological") %>% 
  group_by(GEOID) %>% 
  summarize(disasters=n())

fema_map <- left_join(us_counties, fema_counties_count)

glimpse(fema_map)
```


```{r tigris8, exercise=TRUE}
ggplot(fema_map) +
  geom_sf(aes(fill=disasters)) +
  theme_void()
```

## Styling maps

```{r tigris9, exercise=TRUE}
library(viridis)

ggplot(fema_map) +
  geom_sf(aes(fill=disasters), color=NA) +
  scale_fill_viridis() +
  theme_void()
```


```{r tigris10, exercise=TRUE}
fema_map %>% 
 ggplot() +
  geom_sf(aes(fill=disasters), color=NA) +
  scale_fill_viridis(direction=-1) +
  theme_void()
```


```{r tigris11, warning=F, message=F, exercise=TRUE}
fema_map %>% 
  filter(STUSPS=="LA") %>% 
 ggplot() +
  geom_sf(aes(fill=disasters), color=NA) +
  scale_fill_viridis(direction=-1) +
  theme_void() +
  labs(title="Hurricane declared disasters since 1954")
```

## Small multiples maps

```{r tigris11b, warning=F, message=F, exercise=TRUE}
fema_counties_count_fi <- fema_counties %>%
  filter(incidentType=="Fire") %>% 
  mutate(year=year(incidentBeginDate)) %>% 
  group_by(GEOID, year) %>% 
  summarize(disasters=n())

fema_map_fi <- left_join(us_counties, fema_counties_count_fi)

fema_map_fi %>% 
  filter(STUSPS=="CA") %>% 
  filter(year %in% 2017:2022) %>% 
 ggplot() +
  geom_sf(aes(fill=disasters), color=NA) +
  scale_fill_viridis(direction=-1, na.value="red") +
  facet_wrap(~year) +
  theme_void()
```



```{r tigris12, exercise=TRUE}
fema_counties_count_fi <- fema_counties %>%
  filter(incidentType=="Fire") %>% 
  mutate(year=year(incidentBeginDate)) %>% 
  group_by(GEOID, year) %>% 
  summarize(disasters=n())

fema_map_fi <- left_join(us_counties, fema_counties_count_fi)

fema_map_fi %>% 
  filter(STUSPS=="CA") %>% 
  filter(year %in% 2017:2022) %>% 
 ggplot() +
  geom_sf(data=filter(us_counties, STUSPS=="CA"), fill="gray", color=NA) +
  geom_sf(aes(fill=disasters), color=NA) +
  scale_fill_viridis(direction=-1, na.value="red") +
  facet_wrap(~year) +
  theme_void()
```

## Bivariate maps

*install just cowplot and sf*
`install.packages(c("cowplot", "sf"))`

*install all suggested dependencies*
`install.packages("biscale", dependencies = TRUE)`


```{r tigris13, exercise=TRUE}

disaster_count <- fema_counties %>% 
  filter(incidentType=="Hurricane") %>% 
  mutate(year=year(incidentBeginDate)) %>% 
  filter(year %in% c(2017:2022)) %>% 
  group_by(GEOID) %>% 
  summarize(total=n())

disaster_counties <- left_join(us_counties, disaster_count) %>% 
  select(-NAME)
disaster_counties <- left_join(disaster_counties, county_pop)
  
glimpse(disaster_counties)
```



```{r tigris14, exercise=TRUE}
library(biscale)

bi_df <- bi_class(disaster_counties, x = estimate, y = total, style = "quantile", dim = 3)

glimpse(bi_df)
```


```{r tigris15, exercise=TRUE}
map <- ggplot() +
  geom_sf(data = bi_df, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink", dim = 3) +
  labs(
    title = "Race and Income in St. Louis, MO",
    subtitle = "Gray Pink (GrPink) Palette"
  ) +
  bi_theme()

map
```

```{r tigris16, exercise=TRUE}

legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Higher population ",
                    ylab = "Higher disasters ",
                    size = 8)

legend
```


```{r tigris17, exercise=TRUE}

finalPlot <- ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.82, .2, 0.2, 0.2)

finalPlot

## WHY DOESNT THIS WORK. REPLACE WITH IMAGE :(
```

## save it

```{r, eval=F}
#ggsave(finalPlot, "test.svg", device="svg")
save_plot("test.png", finalPlot, base_height = NULL, base_width = 12)
#install.packages("svglite") to make this work
save_plot("test.svg", finalPlot, base_height = NULL, base_width = 12)

#as a shapefile?
st_write(map, "map.geojson")
st_write(map, "map.shp")

```
