---
title: "Welcome to Advanced Data Journalism: Doing more with R"
author: "Andrew Ba Tran"
output: learnr::tutorial
runtime: shiny_prerendered
# Do not index/display tutorial by setting `private: true`
private: true
description: >
  This is welcome module to see if everything is running smoothly.
---

```{r setup}
packages <- c("tidyverse", "lubridate", "rvest", "httr", "remotes")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos = "https://cran.us.r-project.org")  
}

library(learnr)
```

Welcome to class!

This page checked if the R packages needed for this class were installed on your computer and installed any that were missing.

It's also a way to incorporate each week's lessons and live-coding exercises. 

Every section will be initiated through RStudio in this manner:

`learnr::run_tutorial("welcome", "adjclass")`

or 

`learnr::run_tutorial("class1", "adjclass")`

and so on.

This will allow you to read through the guides and run parts of the R code in the exercises.

The answers needed to answer each week's quiz will be derived from the exercises from these pages.

```{r addition, exercise=TRUE}
1 + 1
```

